Master's Project Journal

### 2019-04-24 thesis/TestingAnalysis.tex
partial corrections
citation and tables
DONEgit add -u!

\section{Chapter Overview}
This chapter contains details of the measurements carried out on networks of various sizes and investigation into the impact of a number of decisions made in the design process. Explanations will be given or analysis made of the results, which will be given in the form of of either plots or tables, depending on which allows for more concise presentation of the data. Both 2x2 and 3x3 network performance analysis was performed with no input delay to the \ac{LF}, which Section \ref{subsection:lf_delay} will show to have a significant performance impact. All tests were carried out, unless otherwise stated, at a centre frequency of $5~\si{\mega\hertz}$.

\section{2x2 Network Performance Comparison}
In order to investigate the impact of the \ac{ADPLL} network on performance, and conversely the impact of each individual \ac{ADPLL} design on the network, switches were used to alter the configuration of the network without requiring re-implementation and a number of captures made at each setting. The three configurations were: each \ac{ADPLL} individually locked to the reference, the network in uni-directional mode and bi-directional mode. The centre frequency or gains were not changed between measurements in order to obtain a good comparison. For \ac{ADPLL} 1 the gains were: $k_p = \frac{1}{32}$ \& $k_i = \frac{1}{256}$. For \ac{ADPLL} 2 and 3 they were: $k_p = \frac{1}{16}$ \& $k_i = \frac{1}{64}$. The

There are two main trends worth highlighting here, with the first of these being as the feedback network becomes more complicated a large degree of skew is introduced, with the magnitude becoming larger in the \acp{ADPLL} further from the reference. This would appear to be attributable to a combination of the time taken for a change in the reference to propagate to the \ac{ADPLL} node in question, and the fineness of period and detection resolutions. This second observation follows from the significant reduction in the skew growth in Designs 2 \& 3 which have finer resolutions.

The other trend, is the variation in \ac{C2C} jitter across the \ac{ADPLL} designs, with coarseness of the \ac{FPGA} clocked \ac{DCO} becoming somewhat obvious. As free \aclp{PLL} (\acsp{PLL}), the clocked \ac{DCO} has almost identical \ac{C2C} jitter and maximum \ac{TIE}, implying that only two period steps are used the majority of the time. As these steps are somewhat further apart than that of the \ac{RO}, at $3.875~\si{\nano\second}$ apart as opposed to an implementation dependant value in the region of $1.1765~\si{\nano\second}$, a much greater jitter is observed. This would confirm the belief that the \ac{FPGA} clocked designs are better suited to operational frequencies where the generated clock period to \ac{FPGA} clock period ratio is significantly lower than the $\frac{3.875~\si{\nano\second}}{200~\si{\nano\second}} = \frac{1}{51.6}$ in this design. Comparing the \ac{C2C} jitter of \ac{ADPLL} 1 \& 2, where the main difference is the choice of \ac{PFD}, the approximate $\frac{1.6~\si{\nano\second}}{0.50~\si{\nano\second}} \approx 3.2$ ratio of observed jitters is almost identical to the $\frac{3.875~\si{\nano\second}}{1.1765~\si{\nano\second}} \approx 3.29$ ratio of the period steps at $5~\si{\mega\hertz}$, implying the \ac{PFD} is at fault for this difference.

The other notable difference in \ac{C2C} jitter, is that the use of an inverter based \ac{TDC}, and its accompanying variation due to implementation differences, has the expected result of additional \ac{C2C} jitter, however, this impact is signfiicantly less than the that of changes in the \ac{RO} period step, and the inverter based \ac{TDC} in combination with the \ac{FPGA} clocked \ac{DCO} would still suffer the same problems as it does in \ac{ADPLL} Design 1 with the clocked \ac{PFD}.

%TODO additional comments with new data?

\begin{table}[!ht]
    \begin{center}
        \begin{footnotesize}
            \setlength{\tabcolsep}{.9\tabcolsep}
            \begin{tabular}{ll|r|r|r|r|r|r|r|r|r|}           
                \cline{3-11}
                && \multicolumn{3}{c|}{Jitter Standard Deviation (ns)} & \multicolumn{3}{c|}{Max. Time Interval Error (ns)} & \multicolumn{3}{c|}{Skew (ns)} \T\\
                \cline{3-11} 
                &&PLL 11&PLL 12&PLL 22    &PLL 11&PLL 12&PLL 22    &PLL 11&PLL 12&PLL 22\T\\
                \hline
                \multicolumn{2}{|l|}{\ac{ADPLL} Design 1}&-&-&-&-&-&-&-&-&-\T\\
                \multicolumn{2}{|r|}{Free PLLs} &1.6051  &1.6146  &1.6251     &1.6985 &1.5833 &1.2712    &0.2700 &0.5817 &0.2491 \T\\
                \multicolumn{2}{|r|}{Uni-dir.}  &1.7693  &1.8463  &1.8862     &7.3060 &7.2565 &10.151    &2.2791 &1.2822 &1.6299 \T\\
                \multicolumn{2}{|r|}{Bi-dir.}   &1.9964  &1.9146  &1.9572     &15.089 &18.283 &19.673    &9.2449 &11.612 &12.052 \T\\
                \hline
                \multicolumn{2}{|l|}{\ac{ADPLL} Design 2}&-&-&-&-&-&-&-&-&-\T\\
                \multicolumn{2}{|r|}{Free PLLs} &0.56222 &0.49224 &0.64124    &5.3039 &7.8626 &5.4634    &&& \T\\
                \multicolumn{2}{|r|}{Uni-dir.}  &0.55040 &0.51284 &0.62796    &5.5312 &7.5359 &7.9697    &&& \T\\
                \multicolumn{2}{|r|}{Bi-dir.}   &0.57056 &0.50403 &0.61644    &11.515 &15.943 &17.428    &&& \T\\
                \hline
                \multicolumn{2}{|l|}{\ac{ADPLL} Design 3}&-&-&-&-&-&-&-&-&-\T\\
                \multicolumn{2}{|r|}{Free PLLs} &0.66757 &0.84477 &0.71308    &5.9142 &6.1353 &5.3939    &-0.36997 &-0.20981 &-1.3344  \T\\
                \multicolumn{2}{|r|}{Uni-dir.}  &0.66934 &0.83894 &0.71271    &5.8779 &7.1746 &6.8914    &-0.36645 &-0.05647 &-0.36178 \T\\
                \multicolumn{2}{|r|}{Bi-dir.}   &0.68372 &0.83193 &0.68570    &7.5705 &9.6292 &9.7715    & 0.68901 & 1.58720 & 1.5553  \T\\
                \hline
                \B                
            \end{tabular}
        \end{footnotesize}
        \caption{2x2 Network performance Comparison.}
        \label{table:2x2perf}
    \end{center}
    \vspace{-0.5cm}
\end{table}
%\FloatBarrier

\section{3x3 Network Performance Comparison}
Subsequent to tests carried out using a 2x2 network, the corresponding tests were carried out using a 3x3 network in order to analyse the impact of a larger network. For these tests the gains were left unchanged but, as the 2x2 and 3x3 networks have different floorplans, the \acp{RO} have changed somewhat. This has especially manifested itself in the exceptional individual \ac{PLL} performance of Design 3.

From analysis of the 2x2 network, it was evident that the best performing design was going to be \ac{ADPLL} 3, as both \ac{ADPLL} 2 \& 3 contain the clocked \ac{PFD} which appears to cause significant skew as the \ac{ADPLL} node indices increase and move further from the external reference. The increased network size appears to have made the lag between reference edge and the rising edges of the individual generated signals worse, which would be expected as the number of clock cycles required to respond to a change in the phase relationship between reference and the furthest node has increased. \ac{ADPLL} 3 suffers least from the greater complication of the network, likely as a result of its period and detector resolution being the finest of the three designs considered.

As \ac{ADPLL} Design 1 suffered significantly more with skew as the node indices increased, so it is no surprise that in the larger network the problem has been amplified, meaning this design is not well suited for use in larger networks at this frequency due to poor resolution. This does not, however, impact the main role of this design method, which is the emulation of designs intended for use in \aclp{ASIC} (\acsp{ASIC}) as the frequency of operation will be much lower than the $5~\si{\mega\hertz}$ used in these tests, which will bring with it a finer resolution.
%TODO fill in this after measurements
\begin{table}[!ht]
    \begin{center}
        \begin{footnotesize}
            \setlength{\tabcolsep}{.9\tabcolsep}
            \begin{tabular}{ll|r|r|r|r|r|r|r|r|r|}           
                \cline{3-11}
                && \multicolumn{3}{c|}{Jitter Standard Deviation (ns)} & \multicolumn{3}{c|}{Max. Time Interval Error (ns)} & \multicolumn{3}{c|}{Skew (ns)} \T\\
                \cline{3-11} 
                &&PLL 11&PLL 12&PLL 22    &PLL 11&PLL 12&PLL 22    &PLL 11&PLL 12&PLL 22\T\\
                \hline
                \multicolumn{2}{|l|}{\ac{ADPLL} Design 1}&-&-&-&-&-&-&-&-&-\T\\
                \multicolumn{2}{|r|}{Free PLLs} &&& &&& &&& \T\\
                \multicolumn{2}{|r|}{Uni-dir.}  &&& &&& &&& \T\\
                \multicolumn{2}{|r|}{Bi-dir.}   &&& &&& &&& \T\\
                \hline
                \multicolumn{2}{|l|}{\ac{ADPLL} Design 2}&-&-&-&-&-&-&-&-&-\T\\
                \multicolumn{2}{|r|}{Free PLLs} &0.47167 &0.55012 &0.36251    &5.9649 &6.6965 &5.5010    &3.4045&3.2369&2.3108 \T\\
                \multicolumn{2}{|r|}{Uni-dir.}  &0.49292 &0.55586 &0.39054    &5.958  &8.9250 &11.622    &3.5052&6.9860&8.874 \T\\
                \multicolumn{2}{|r|}{Bi-dir.}   &0.43263 &0.48053 &0.33778    &18.867 &35.973 &38.773    &15.193&28.965&30.841 \T\\
                \hline
                \multicolumn{2}{|l|}{\ac{ADPLL} Design 3}&-&-&-&-&-&-&-&-&-\T\\
                \multicolumn{2}{|r|}{Free PLLs} &0.89042 &0.64889 &0.57405    &4.0634 &3.2467 &2.3141    &1.8373&1.1006&0.39334 \T\\
                \multicolumn{2}{|r|}{Uni-dir.}  &0.93580 &0.93353 &0.71361    &4.2675 &5.8193 &6.4139    &1.8605&2.3172&3.1463  \T\\
                \multicolumn{2}{|r|}{Bi-dir.}   &0.97717 &0.85131 &0.63863    &5.6439 &8.4856 &9.4467    &2.7129&4.1074&5.1717  \T\\
                \hline
                \B                
            \end{tabular}
        \end{footnotesize}
        \caption{3x3 Network performance Comparison.}
        \label{table:3x3perf}
    \end{center}
    \vspace{-0.5cm}
\end{table}
%\FloatBarrier

@@

\subsection{Impact of Gain Variation}
The gains chosen for use in the \acl{LF} have significant impact on the performance of the system, as the \ac{LF} governs the response of the \ac{ADPLL} to a given phase error. To large a proportional gain will result in overly aggressive overcompensation, thus negatively affecting the jitter. Eventually the violence of the jitter will result in a loss of lock. However, if the gain is too small, the frequency or period adjustment will not be sufficiently large to cause the \ac{ADPLL} to lock. Similarly, appropriate values of $k_i$ are required, and the relationship between proportional and integral gains must satisfy the conditions set out by Koskin \cite{koskin2018generation}.

@@

Figure \ref{fig:gain_sweep} contains a sweep across the possible range of integral gains for four different proportional gains. This test was carried out with a 3x3 network of \ac{ADPLL} Design 3, each locked to the external reference at $5~\si{\mega\hertz}$, and the generated waveform from the \acp{ADPLL} at locations ``11'', ``22'' and  ``33'' were captured. It can clearly be observed that there are two distinct regions in the graph, with jitter values of less than $1~\si{\nano\second}$ contrasting sharply with the points with jitter in the region of $10~\si{\nano\second}$. These distinct sections of the plot represent locked and unlocked operation respectively. Zooming in on the low jitter region in Figure \ref{fig:gain_sweep_zoom}, the approximate loss of lock thresholds can be compared with the theoretical relationship of $k_i = 0.5\times k_p$. The approximated loss of lock is determined as gain setting that caused the first \ac{ADPLL} to lose lock.
\begin{table}[!h]
    \begin{center}
        \begin{tabular}{lrr}
            \multicolumn{1}{c}{$k_p$} & \multicolumn{1}{c}{$k_i$} & \multicolumn{1}{c}{Ratio} \T\B\\
            \hline
            $\frac{1}{16}$            & $\frac{5}{64}$            & $0.8$ \T\B\\
            $\frac{2}{16}$            & $\frac{11}{64}$           & $0.72$ \T\B\\
            $\frac{3}{16}$            & $\frac{13}{64}$           & $0.92$ \T\B\\
            $\frac{4}{16}$            & $\frac{15}{64}$           & $1.07$ \T\B
        \end{tabular}
    \end{center}
    \vspace{-0.5cm}
    \caption[Loss of Lock Gains]{Loss of Lock Gains.}
    \label{table:lolgains}
\end{table}

Despite the relatively small number of proportional gains, it is clear that the relationship set out by Koskin does not hold in this system, but rather a scaled version of it.  This difference is attributable to a difference in \ac{LF} structure. In his simulations the \ac{LF} design contains a register forming an input buffer, absent in the \ac{LF} used for these tests. The presence of this register, clocked on the generated signal, causes the \ac{ADPLL} to react more slowly to changes in the phase difference, thus increasing the jitter. The removal of this register causes a change in the relationship between proportional and integral gains, a change that will be investigated further in a later section.

\subsection{Distribution of Period Steps}
To add context to the jitter numbers produced for each oscillator previously, the variation of period steps in the \ac{RO}, and delays in the \ac{TDL} due to implementation should be examined. The best way to visualise this distribution is through a histogram, placing periods of in a certain range into the same bin. This can be visualised effectively using Matlab's \texttt{histplot}. In order to obtain data, three \acp{ADPLL} implementing both the \ac{FPGA} clocked and inverter based \acp{DCO} were locked to an external reference signal at $5~\si{\mega\hertz}$, and the period of each clock cycle measured.
\begin{figure}[h]%
    \centering
    \subfloat[Inverter propagation delay based \ac{RO}.]{{\includegraphics[width=0.5\textwidth]{../conf_paper/distrib_ring} }}%
    \subfloat[\ac{FPGA} clocked \ac{DCO}.]{{\includegraphics[width=0.5\textwidth]{../conf_paper/distrib_pa} }}%
    \caption[Example distribution of periods]{Example distribution of periods.}    
    \label{fig:dists}
\end{figure}
Figure \ref{fig:dists} contains the output of \texttt{histfit} for each of the three oscillators superimposed in one plot for each \ac{DCO} type. As the \ac{RO} has significantly finer period resolution, the period steps would be overlaid on each-other thereby making the plot difficult to read. In order to avoid this behaviour, the \ac{RO}'s mapping of control code to inverters to remove was altered from twice the control code to 16 times the control code. This results in a more pronounced separation of period steps while being indicative of the variation due to implementation.
The plots paint a a picture of the unrealistic nature of the \ac{FPGA} clocked design in comparison to its inverter based counterpart, as each oscillator will have identical intrinsic period steps thus eliminating any jitter that maybe be seen in the system due to misalignment.
This effects the suitability of the FPGA clocked \ac{DCO} for the improvement of simulation models, or the examination of how a novel block may impact the behaviour of the network. If this analysis was carried out for the period steps, similar behaviour would be observed.

>>>>

\section{Chapter Overview}
This chapter contains details of the measurements carried out on networks of various sizes in addition to investigation into the impact of a number of decisions made in the design process. Explanations will be given, or analysis made, of the results, which will be either in the form of plots or tables, depending on which allows for more concise presentation of the data. Both 2x2 and 3x3 network performance analysis was performed with no input delay to the \ac{LF}, which Section \ref{subsection:lf_delay} will show to have had a significant performance impact. Unless otherwise stated, all tests were carried out at a centre frequency of $5~\si{\mega\hertz}$.

\section{2x2 Network Performance Comparison}
In order to investigate the impact of the \ac{ADPLL} network on performance, and conversely the impact of each individual \ac{ADPLL} design on the network, switches were used to alter the configuration of the network without requiring re-implementation. A number of captures were made at each setting. The three configurations were: each \ac{ADPLL} individually locked to the reference, the network in uni-directional mode and in bi-directional mode. In order to obtain a good comparison, the \ac{LF} gains were not changed between measurements. For \ac{ADPLL} 1 the gains were: $k_p = \frac{1}{32}$ \& $k_i = \frac{1}{256}$. For \ac{ADPLL} 2 and 3 they were: $k_p = \frac{1}{16}$ \& $k_i = \frac{1}{64}$. \ac{ADPLL} requires a smaller gain due to the greater frequency sensitivity.

\begin{table}[!ht]
    \begin{center}
        \begin{footnotesize}
            \setlength{\tabcolsep}{.9\tabcolsep}
            \begin{tabular}{ll|r|r|r|r|r|r|r|r|r|}           
                \cline{3-11}
                && \multicolumn{3}{c|}{Jitter Standard Deviation (ns)} & \multicolumn{3}{c|}{Max. Time Interval Error (ns)} & \multicolumn{3}{c|}{Skew (ns)} \T\\
                \cline{3-11} 
                &&PLL 11&PLL 12&PLL 22    &PLL 11&PLL 12&PLL 22    &PLL 11&PLL 12&PLL 22\T\\
                \hline
                \multicolumn{2}{|l|}{\ac{ADPLL} Design 1}&-&-&-&-&-&-&-&-&-\T\\
                \multicolumn{2}{|r|}{Free PLLs} &1.6051  &1.6146  &1.6251     &1.6985 &1.5833 &1.2712    &0.2700 &0.5817 &0.2491 \T\\
                \multicolumn{2}{|r|}{Uni-dir.}  &1.7693  &1.8463  &1.8862     &7.3060 &7.2565 &10.151    &2.2791 &1.2822 &1.6299 \T\\
                \multicolumn{2}{|r|}{Bi-dir.}   &1.9964  &1.9146  &1.9572     &15.089 &18.283 &19.673    &9.2449 &11.612 &12.052 \T\\
                \hline
                \multicolumn{2}{|l|}{\ac{ADPLL} Design 2}&-&-&-&-&-&-&-&-&-\T\\
                \multicolumn{2}{|r|}{Free PLLs} &0.66275 &0.62108 &0.63211    &5.7432 &5.4399 &5.2317    &0.00768 &-0.1681 &0.5383  \T\\
                \multicolumn{2}{|r|}{Uni-dir.}  &0.65990 &0.63995 &0.67278    &5.6926 &6.9422 &8.9609    &0.00663 &-0.3453 &0.76741 \T\\
                \multicolumn{2}{|r|}{Bi-dir.}   &0.66308 &0.66233 &0.67583    &12.222 &16.719 &19.176    &5.0799  & 7.2752 &8.7174  \T\\
                \hline
                \multicolumn{2}{|l|}{\ac{ADPLL} Design 3}&-&-&-&-&-&-&-&-&-\T\\
                \multicolumn{2}{|r|}{Free PLLs} &0.66757 &0.84477 &0.71308    &5.9142 &6.1353 &5.3939    &-0.36997 &-0.20981 &-1.3344  \T\\
                \multicolumn{2}{|r|}{Uni-dir.}  &0.66934 &0.83894 &0.71271    &5.8779 &7.1746 &6.8914    &-0.36645 &-0.05647 &-0.36178 \T\\
                \multicolumn{2}{|r|}{Bi-dir.}   &0.68372 &0.83193 &0.68570    &7.5705 &9.6292 &9.7715    & 0.68901 & 1.58720 & 1.5553  \T\\
                \hline
                \B                
            \end{tabular}
        \end{footnotesize}
        \caption{2x2 Network Performance Comparison.}
        \label{table:2x2perf}
    \end{center}
    \vspace{-0.5cm}
\end{table}
There are two main trends worth highlighting here, the first of these being as the feedback network becomes more complicated a large degree of skew is introduced, with the magnitude becoming larger in the \acp{ADPLL} further from the reference. This would appear to be attributable to a combination of the time taken for a change in the reference to propagate to the \ac{ADPLL} node in question, and the fineness of period and detection resolutions. This second observation follows from the significant reduction in the skew growth in Designs 2 \& 3 which have finer resolutions.

The other trend, is the variation in \ac{C2C} jitter across the \ac{ADPLL} designs, with coarseness of the \ac{FPGA} clocked \ac{DCO} becoming somewhat obvious. As free \aclp{PLL} (\acsp{PLL}), the clocked \ac{DCO} has almost identical \ac{C2C} jitter and maximum \ac{TIE}, implying that only two period steps are used the majority of the time. As, at $3.875~\si{\nano\second}$, these steps are somewhat further apart than that of the \ac{RO} (an implementation dependant value in the region of $1.1765~\si{\nano\second}$), a much greater jitter is observed. This would confirm the belief that the \ac{FPGA} clocked designs are better suited to operational frequencies where the generated clock period is multiple orders of magnitude longer than the \ac{FPGA} clock period. In this design, the ratio works out as $\frac{3.875~\si{\nano\second}}{200~\si{\nano\second}} = \frac{1}{51.6}$, which is insufficient. Comparing the \ac{C2C} jitter of \ac{ADPLL} 1 \& 2, where the main difference is the choice of \ac{RO}, the approximate $\frac{1.6~\si{\nano\second}}{0.50~\si{\nano\second}} \approx 3.2$ ratio of observed jitters is almost identical to the $\frac{3.875~\si{\nano\second}}{1.1765~\si{\nano\second}} \approx 3.29$ ratio of the period steps at $5~\si{\mega\hertz}$, implying the \ac{PFD} is at fault for this difference.

The other notable difference in \ac{C2C} jitter, is that the use of an inverter based \ac{TDC}, and its accompanying variation due to implementation differences, has the expected result of additional \ac{C2C} jitter. However, this impact is significantly less than that of changes in the \ac{RO} period step. The combination of inverter based \ac{TDC} in combination with the \ac{FPGA} clocked \ac{DCO} would likely be wasted, as the \ac{DCO} would still suffer the same period distribution issues it does in \ac{ADPLL} Design 1, with the clocked \ac{PFD}.

Overall, the best performer is Design 3, as although the \ac{C2C} jitter is worse that of Design 2, it exhibits the least skew in bi-directional mode. This would imply that period and detection resolution are the key attribute guiding network performance. \ac{ADPLL} 3's performance confirms it's status as the ideal choice for the testing of new control or performance improvement techniques, or comparison with theoretical models.

\section{3x3 Network Performance Comparison}
Subsequent to tests carried out using a 2x2 network, the corresponding tests were carried out using a 3x3 network in order to analyse the impact of a larger network. For these tests the gains were left unchanged but, as the 2x2 and 3x3 networks have different floorplans, the \ac{RO} layouts have changed somewhat. This layout change has especially manifested itself in the exceptional individual \ac{PLL} performance of Design 3.

\begin{table}[!ht]
    \begin{center}
        \begin{footnotesize}
            \setlength{\tabcolsep}{.9\tabcolsep}
            \begin{tabular}{ll|r|r|r|r|r|r|r|r|r|}           
                \cline{3-11}
                && \multicolumn{3}{c|}{Jitter Standard Deviation (ns)} & \multicolumn{3}{c|}{Max. Time Interval Error (ns)} & \multicolumn{3}{c|}{Skew (ns)} \T\\
                \cline{3-11} 
                &&PLL 11&PLL 22&PLL 33    &PLL 11&PLL 22&PLL 33    &PLL 11&PLL 22&PLL 33 \T\\
                \hline
                \multicolumn{2}{|l|}{\ac{ADPLL} Design 1}&-&-&-&-&-&-&-&-&-\T\\
                \multicolumn{2}{|r|}{Free PLLs} &1.8956  &1.8425  &1.7594     &5.7868 &5.8407 &6.2537    &1.8481 &1.5976 &2.3251 \T\\
                \multicolumn{2}{|r|}{Uni-dir.}  &1.7836  &1.8206  &1.8827     &5.5863 &6.522  &9.4008    &1.7697 &1.0359 &1.5503 \T\\
                \multicolumn{2}{|r|}{Bi-dir.}   &1.9049  &1.9205  &1.9441     &22.980 &42.771 &47.441    &17.268 &33.739 &35.989 \T\\
                \hline
                \multicolumn{2}{|l|}{\ac{ADPLL} Design 2}&-&-&-&-&-&-&-&-&-\T\\
                \multicolumn{2}{|r|}{Free PLLs} &0.47167 &0.55012 &0.36251    &5.9649 &6.6965 &5.5010    &3.4045&3.2369&2.3108   \T\\
                \multicolumn{2}{|r|}{Uni-dir.}  &0.49292 &0.55586 &0.39054    &5.958  &8.9250 &11.622    &3.5052&6.9860&8.874    \T\\
                \multicolumn{2}{|r|}{Bi-dir.}   &0.43263 &0.48053 &0.33778    &18.867 &35.973 &38.773    &15.193&28.965&30.841   \T\\
                \hline
                \multicolumn{2}{|l|}{\ac{ADPLL} Design 3}&-&-&-&-&-&-&-&-&-\T\\
                \multicolumn{2}{|r|}{Free PLLs} &0.89042 &0.64889 &0.57405    &4.0634 &3.2467 &2.3141    &1.8373&1.1006&0.39334  \T\\
                \multicolumn{2}{|r|}{Uni-dir.}  &0.93580 &0.93353 &0.71361    &4.2675 &5.8193 &6.4139    &1.8605&2.3172&3.1463   \T\\
                \multicolumn{2}{|r|}{Bi-dir.}   &0.97717 &0.85131 &0.63863    &5.6439 &8.4856 &9.4467    &2.7129&4.1074&5.1717   \T\\
                \hline
                \B                
            \end{tabular}
        \end{footnotesize}
        \caption{3x3 Network Performance Comparison.}
        \label{table:3x3perf}
    \end{center}
    \vspace{-0.5cm}
\end{table}
The results of the 3x3 network testing are largely similar to those from before, with the same observations applying. From analysis of the 2x2 network, it was evident that the best performing design was going to be \ac{ADPLL} 3, as both \ac{ADPLL} 2 \& 3 contain the clocked \ac{PFD} which appears to cause significant skew as the \acp{ADPLL} move further from the external reference. The increased network size appears to have made the lag between reference edge and the rising edges of the individual generated signals worse, which would be expected as the number of clock cycles required to respond to a change in the phase relationship between reference and the furthest node has increased. \ac{ADPLL} 3 suffers least from the greater complication of the network, likely as a result of its period and detector resolution being the finest of the three designs considered.

As \ac{ADPLL} Design 1 suffered significantly more with skew as the node indices increased, it is no surprise that in the larger network the problem has been amplified, meaning this design is not well suited for use in larger networks at this frequency, due to poor resolution. This does not, however, impact the main role of this design method, which is the emulation of designs intended for use in \aclp{ASIC} (\acsp{ASIC}) at scaled frequencies. As the frequency of operation should be much lower than the $5~\si{\mega\hertz}$ used in these tests, the finer resolution, relative to the generated period, should alleviate this skew.

@@

\subsection{Impact of Gain Variation}
The gains chosen for use in the \acl{LF} have significant impact on the performance of the system, as the \ac{LF} governs the response of the \ac{ADPLL} to a given phase error. Too large a proportional gain will result in compensation, thus negatively affecting the jitter. Eventually the violence of the jitter will result in a loss of lock. However, if the gain is too small, the frequency or period adjustment will not be sufficiently large to cause the \ac{ADPLL} to lock. Similarly, appropriate values of $k_i$ are required, and the relationship between proportional and integral gains must satisfy the conditions set out by Koskin \cite{koskin2018generation}.

@@

Figure \ref{fig:gain_sweep} contains a sweep across the possible range of integral gains for four different proportional gains. This test was carried out with a 3x3 network of \ac{ADPLL} Design 3, each locked to the external reference at $5~\si{\mega\hertz}$. Only the generated waveforms from the \acp{ADPLL} at locations ``11'', ``22'' and  ``33'' were captured, alongside the external reference. It can be clearly observed that there are two distinct regions in the graph, with jitter values of less than $1~\si{\nano\second}$ contrasting sharply with the section where jitter is in the region of $10~\si{\nano\second}$. These distinct sections of the plot represent locked and unlocked operation respectively. Zooming in on the low jitter region in Figure \ref{fig:gain_sweep_zoom}, the approximate loss of lock thresholds can be compared with the theoretical relationship of $k_i = 0.5\times k_p$. The approximated loss of lock is determined as gain setting that caused the first \ac{ADPLL} to lose lock.
\begin{table}[!h]
    \begin{center}
        \begin{tabular}{lrr}
            \multicolumn{1}{c}{$k_p$} & \multicolumn{1}{c}{$k_i$} & \multicolumn{1}{c}{Ratio} \T\B\\
            \hline
            $\frac{1}{16}$            & $\frac{5}{64}$            & $0.8$ \T\B\\
            $\frac{2}{16}$            & $\frac{11}{64}$           & $0.72$ \T\B\\
            $\frac{3}{16}$            & $\frac{13}{64}$           & $0.92$ \T\B\\
            $\frac{4}{16}$            & $\frac{15}{64}$           & $1.07$ \T\B
        \end{tabular}
    \end{center}
    \vspace{-0.5cm}
    \caption[Loss of Lock Gains]{Loss of Lock Gains.}
    \label{table:lolgains}
\end{table}

Despite the relatively small number of proportional gains, it is clear that the exact relationship set out by Koskin does not hold in this system, but rather a scaled version of it does.  This difference is attributable to a modification in the structure of the \ac{LF}. In his simulations, the \ac{LF} design contains a register forming an input buffer, absent in the \ac{LF} used for these tests. The presence of this register, clocked on the generated signal, causes the \ac{ADPLL} to react more slowly to changes in the phase difference, thus increasing the jitter. This register was removed as it is not required to meet timing constraints in the \acp{ADPLL} used in this project. This removal allows the system to respond a cycle faster to a change in the phase relationships, and the impact of this change will be investigated further in a later section.

\subsection{Distribution of Period Steps}
To add context to the jitter numbers produced for each oscillator previously, the variation of period steps in the \ac{RO}, and delays in the \ac{TDL} due to implementation should be examined. The best way to visualise this distribution is through a histogram, placing periods within in a certain range into the same bin. This can be visualised effectively using Matlab's \texttt{histplot}. In order to obtain data, three instances of \acp{ADPLL} implementing each of the \ac{FPGA} clocked and inverter based \acp{DCO} were locked to an external reference signal at $5~\si{\mega\hertz}$. A capture was taken of each generated signal, the period of each clock cycle measured based on their rising edges.

\begin{figure}[h]%
    \centering
    \subfloat[Inverter propagation delay based \ac{RO}.]{{\includegraphics[width=0.5\textwidth]{../conf_paper/distrib_ring} }}%
    \subfloat[\ac{FPGA} clocked \ac{DCO}.]{{\includegraphics[width=0.5\textwidth]{../conf_paper/distrib_pa} }}%
    \caption[Example distribution of periods]{Example distribution of periods.}    
    \label{fig:dists}
\end{figure}
Figure \ref{fig:dists} contains the output of \texttt{histfit} for each of the three oscillators, superimposed in one plot for each \ac{DCO} type. As the \ac{RO} has significantly finer period resolution, the period steps would be overlaid on each-other, thereby making the plot difficult to read. In order to avoid this behaviour, the \ac{RO}'s mapping of control code to inverters was altered, from twice the control code to 16 times the control code. This resulted in a more pronounced separation of period steps while being indicative of the variation due to implementation.
The plots paint a a picture of the unrealistic nature of the \ac{FPGA} clocked design in comparison to its inverter based counterpart, as each \ac{DCO} has identical intrinsic period steps thus eliminating any jitter that may be seen in the system due to misalignment.
This effects the suitability of the FPGA clocked \ac{DCO} for the improvement of simulation models, or the examination of how a novel block may impact the behaviour of the network. If this analysis was carried out for the period steps, similar behaviour would be observed.

\subsection{\ac{FPGA} Clocked \ac{DCO} Width Variation}
When introducing the design of the \ac{DCO} for \ac{ADPLL} 1, the minimum possible accumulator width was chosen. From the equation used to calculate the frequency step size, it may have seemed that by doubling the bias point and increasing the width of the counter to 13 bits a finer resolution could be achieved.
\begin{align}
f_{step} &= \frac{f_{FPGA}}{2^{12}} = 62.988~\si{\kilo\hertz} \\
f_{step} &= \frac{f_{FPGA}}{2^{13}} = 31.494~\si{\kilo\hertz}
\end{align} Both the following measurements were carried out using identical loop filter gains, network configurations and reference signal. While in uni-directional mode some minor difference is observed in the maximum \ac{TIE}, but overall no statistically relevant changes were seen.

As Table \ref{table:accum_width} shows, while the resolution may have improved, the jitter has not. This occurs as all signals are synchronised to the \ac{FPGA} clock, which has a $3.875~\si{\nano\second}$ period. This quantises the output waveform to the \ac{FGPA} clock, and sets the minimum bound on the jitter. Accordingly, the most suitable counter width is that which maximises the bit width, and thus achieves the best frequency resolution, with the constraint that, at the operating frequency, the a period resolution is no finer than the \ac{FPGA} clock period. In this case, there is no bit width that satisfies both constraints, so the latter is relaxed to get a width of $12$.
%TODO explanation for this?
\begin{table}[!ht]
    \begin{center}
        \begin{footnotesize}
            \setlength{\tabcolsep}{.9\tabcolsep}
            \begin{tabular}{ll|r|r|r|r|r|r|r|r|r|}           
                \cline{3-11}
                && \multicolumn{3}{c|}{Jitter Standard Deviation (ns)} & \multicolumn{3}{c|}{Max. Time Interval Error (ns)} & \multicolumn{3}{c|}{Skew (ns)} \T\\
                \cline{3-11} 
                &&PLL 11&PLL 12&PLL 22    &PLL 11&PLL 12&PLL 22    &PLL 11&PLL 12&PLL 22\T\\
                \hline
                \multicolumn{2}{|l|}{Accum. Width 12}&-&-&-&-&-&-&-&-&-\T\\
                \multicolumn{2}{|r|}{Uni-dir.}  &1.7693 &1.8463 &1.8862    &7.3060  &7.2565  &10.151    &2.2791 &1.2822  &1.6299  \T\\
                \multicolumn{2}{|r|}{Bi-dir.}   &1.9964 &1.9146 &1.9572    &15.0890 &18.2830 &19.673    &9.2449 &11.6120 &12.0520 \T\\
                \hline
                \multicolumn{2}{|l|}{Accum. Width 13}&-&-&-&-&-&-&-&-&-\T\\
                \multicolumn{2}{|r|}{Uni-dir.}  &1.7946 &1.8161 &1.9048    &6.8614  &7.1495 &9.371      &2.2257 &1.6328  &1.3778  \T\\
                \multicolumn{2}{|r|}{Bi-dir.}   &2.0016 &1.9217 &1.9605    &14.9580 &18.938 &19.510     &9.1678 &12.0000 &11.7690 \T\\
                \hline
                \B
            \end{tabular}
        \end{footnotesize}
        \caption{\ac{DCO} Accumulator Width Comparison.}
        \label{table:accum_width}
    \end{center}
    \vspace{-0.5cm}
\end{table}
%\FloatBarrier

\subsection{\ac{LF} Input Delay Register}\label{subsection:lf_delay}
Previously, when comparing the loss of lock gain relationship to that proposed by Koskin \textit{et al}, it was mentioned that the \ac{LF} used in this project does not have an input delay register. The lack of agreement between the theory and experimental results was attributed to the input delay register, or lack thereof. In order to provide a basis for this claim, the performance of a network in 2x2 configuration was analysed and the results of this analysis are presented in Table \ref{table:lf_delay}. While the \ac{C2C} jitter is statistically similar, there is significant difference to be observed in the \ac{TIE}, where an approximately $50\%$ increase can be observed, measurement for measurement. The absolute maximum \ac{TIE} is significantly impacted by the presence of any skew, as skew is the average value of this quantity, it is important to note that the skew has not meaningfully changed. As skew has not changed, it can be determined that jitter is the source of the increased \ac{TIE}.
